Predicting I/O-performance in HPC using Artificial Neural Networks

### 2016-07-07 Paper/paper.tex
Kommentare
Nai

%\classify{MSC?}
\author{Fabian Schmidt\footnote{\label{susu}Universität Hamburg, E-Mail:\url{2schmid@informatik.uni-hamburg.de}} \and Julian M. Kunkel\footnote{German Climate Computing Center (DKRZ)}}

@@

\keywords{file systems, performance, predicting file access times, artificial neural networks}%TODO Gute keywords?
\end{abstract}

@@

Because of the strong linear correlation between access time and access size, the problem seems to fit linear models.
However, our results show that the relation of file access parameters to access time is not sufficiently represented by linear models.
ANNs achieve significantly better results than linear models.
Our analysis suggests that the I/O-path used by the storage system significantly influences the file access time.
Therefore it becomes key for a good model of access times to derive knowledge about I/O-paths.
Unfortunately I/O-paths are difficult to deal with, as it is unknown which path was used for a file access.

This paper is organized as follows:
Related work is provided in Section\,\ref{sec:related}.
In Section\,\ref{modeling_access_times}, we explain our analysis of file access times. 
Firstly, we develop a simplified model of the I/O-path. 
Next, a method for approximating I/O-paths with derived classes from the error of a prediction model is proposed. 
At last, we introduce a set of models for access time prediction.
Afterwards in Section\,\ref{evaluation}, we evaluate our analysis of the storage system by measuring access times in various test series, studying the obtained access times and the predictions of our models to them.
Finally, Section\,\ref{conclusion} summarizes the paper and suggests future work.

@@

Generally, storage systems are modeled in two different ways for access time prediction; either using white-box modeling or black-box modeling~\cite{Crume:2013:FML:2538542.2538561}.
\begin{itemize}
	\item \textbf{White-box modeling}: The storage system itself is simulated. 
	Details of hardware components like rotation speed of the magnetic disk in a hard drive are considered. 
	The processing of a file access can then be simulated in the model and the resulting access time is then used as prediction for the actual system.
	Processing and resulting performance can be analyzed in detail on the model.
	\item \textbf{Black-box modeling}: The model abstracts from the real storage system. 
	System performance is approximated without considering the causes. %Zustandekommen
	This procedure corresponds to an emulation of the storage system.
	In contrast to the white-box model, processing of file accesses can't be analyzed on the model itself.
\end{itemize}
These two ways of modeling are fundamentally different and have to be differentiated.

\subsection{White-box-modeling versus black-box-modeling}
For the in-depth analysis of reasoning for behavior of a storage system, a white-box-model is desirable.
On the one hand, the modeled system is represented in the model and, thus, can be examined. 
On the other hand, these models can be very precise if modeled correctly \cite{Ruemmler94anintroduction}.
There are important limitations of white-box modeling, however. 
For every system an individual model has to be created and a model becomes quite intricate for single hard drives already \cite{Crume:2013:FML:2538542.2538561}.
To approach the complexity of white-box modeling, Ruemmler and Wilkes analyzed the relevance of different hard drive components for the model deviation, to save effort for insignificant parts \cite{Ruemmler94anintroduction}.
However, white-box modeling is usually used for simple systems like a single hard drive. For these hard drives white-box-modeling is already very demanding, hence for the complex parallel storage system of a supercomputer it's not a feasible approach \cite{DBLP:conf/npc/ZhangLZJC10}.\\

Application of black-box-modeling is easier and more flexible as it's independent from the individual system.
Stochastic approaches coupled with data mining methods are mostly used for black-box modeling; for example, a combination of regression trees support vector regression \cite{Dai:2012:SDP:2477169.2477214}, or selective bagging classification and regression trees \cite{DBLP:conf/npc/ZhangLZJC10}.

\subsection{Prediction of I/O-performance with ANNs}
Computability of ANNs was researched by Rojas \cite{Rojas:1996:NNS:235222} and Cybenko \cite{cybenko:mcss}; they demonstrated possibility of modeling non-linear systems. 
Cybenko also proved the \textit{universal approximation theorem}, which states that feed-forward networks with sufficient complexity can approximate any continuous functions on compact subsets of $\mathbb{R}^n$.
Therefore ANNs should be sufficient for predicting I/O-performance as long as there are no contradictions in behavior. 
However, no statement can be made about the necessary structure of the network for this task.\\
Crume et al. developed a method using ANNs which exploits periodic patterns in sequences of file access times \cite{Crume:2013:FML:2538542.2538561}.
A Fourier analysis is used to determine the most important frequencies, which are then used as input information for the ANNs.
In a following publication \cite{crumelatent} they move away from Fourier analysis, but then use additional sine waves as input for the ANNs.
This seems to be a promising approach for predicting access times of single hard drives, where the rotational characteristics of the magnetic disk are an essential consideration. 
It is, however, difficult to extrapolate behavior from a single disk to a distributed storage system consisting of thousands of disks that utilizes several optimization mechanisms.

\subsection{I/O-performance prediction in HPC}
Performance analysis in HPC is a important task to examine the system for improvements in usage.
For instance Liu et al. simulated scheduling algorithms for research \cite{liu2011towards}. 
The simulation used the white-box modeling tool DiskSim for prediction of occurring file accesses \cite{Bucy08thedisksim}.\\
There are a few simulators for parallel storage systems, for example, CODES\,\cite{cope2011codes} utilizes a scalable infrastructure to investigate relevant research issues such as the importance of burst-buffers.
PIOSimHD is a simulator able to replay (MPI) application traces on a generic I/O system \cite{kunkel2013simulating}.

@@

\section{Modeling I/O Access Times}
\label{modeling_access_times}
\subsection{Characteristics of the Data}
We used a synthetic benchmark in which identical measurements for random or sequential I/O are performed and the access time is measured for each operation individually.
The resulting timings are stored together with file access parameters in a file that is then used to build and validate the models.
This approach allows for an in-depth analysis of system behavior for different use cases from the perspective of a single client.
The stored parameters are:
\begin{itemize}
%	\item \textbf{File-ID}: Identifies the invoked file. Brauchen wir ja nicht für die Analyse
	\item \textbf{Access size}: Number of bytes to read or write.
	\item \textbf{Access type}: Differentiates reading and writing.
\end{itemize}
Directly measurable or derivable attributes are:
\begin{itemize}
	\item \textbf{Offset}: Distance of file beginning to the starting point of access.
	\item \textbf{Delta-Offset}: Can be calculated for file access $i$ as Offset[$i$] - (Offset[i-1] + access size[i-1]).
	\item \textbf{Access time}: Time for performing the I/O.
\end{itemize}
Knowledge of internal aspects of the system about the current system utilization or about the storage media that have to be addressed is not utilized in the model.

\subsection{Model of the I/O-path}
The internal processing of a file access in the storage system can be viewed using the I/O-path, which is the path from the invoking processor to the storage medium that contains/will store the data (larger accesses may be distributed across multiple servers). 
The data read/written it is passed through the levels of the memory hierarchy.
The resulting access time depends on the depth of this I/O-path because storage media further along the path are increasingly slower.
While the first levels in memory hierarchy (CPU caches) are extremely fast, the main memory is already slower; the same applies for the step into the parallel storage system, that is deployed as servers connected via network to the computer nodes.
%The trade-off between memory capacity and performance can be visualized with the memory hierarchy as in figure %\ref{src/2000px-ComputerMemoryHierarchy.png}.
%\fig{width=.7\textwidth}{src/2000px-ComputerMemoryHierarchy.png}{Memory hierarchy.}
Reading file accesses are more severely affected by varying depths of I/O-paths than writing file accesses, which can be deferred using write-behind and propagated lazily to the disk drives.

@@

At a first thought, one could come up with the idea of simply clustering the data seen in Figure~\ref{src/plot_SizeSorted_log_read_seq.png}, because the groups of measurements with varying I/O-paths might be correctly differentiated by it. 
This, however, is not a good approach, one would have to cluster for every set of measurements with identical access parameters individually as access time is strongly dependent on the access size. 
In a real application scenario, a large number of different access parameters with only very few instances occur, which makes this approach unfeasible.\\
Another idea might be then to use the difference of an arbitrary value to the measured file access time as parameter for the clustering. 
Such a value would not be appropriate for all magnitudes of access time occurring in the data.
I/O-paths that differ by little might not be found with this approach, because the gap between them doesn't matter at the scale of the chosen value.

Our approach uses the residues of a model like linear regression that can't differentiate between measurements with equal access parameter values.
Basically the linear regression predicts an average access time for the measurements of any given size.
The deviation of an individual measurement for these access parameters to the predicted average value is then characteristic for the I/O-path.
Clustering the residues with a k-Means algorithm allows us then to assign each measurement to a specific cluster, which is the approximation of its I/O-path.
We call the retrieved clusters \textbf{error classes}.
We use the absolute error of linear regression as clustering parameter so that one cluster can, for example, represent an positive deviation of 1 millisecond.
Ideally, it might be the case that file accesses independent of the access parameters that took 1 millisecond longer than usual are processed on the same I/O-path.
In that case error classes are directly corresponding to I/O-paths. 
However, this has to be investigated.

@@

\subsection{Models}
Access times of file accesses can be predicted using various models.
While we evaluated several models for this paper, only a few relevant models are described; more models are described in \cite{VVEIHUDVVN15}. 
% TODO: Soll ich meine Arbeit noch woanders zitieren?
Each model seeks to correlate its known file access parameters to the corresponding access time. 
Linear regression is used as a baseline model with a simple mapping of access size to access time.
Additionally, three models with different input information utilizing ANNs are used:
\begin{itemize}
	\item The most simple ANN-model only received information about access sizes, delta-offsets and access types as input.
	\item A second ANN-model, called \textbf{ema-model}, received additionally to the parameters of the previous model information about the past data throughputs of the system, which can be used to exploit time dependencies of the I/O-performance.
	\item The third ANN-model, called \textbf{error-class-model}, received error classes in addition to the access parameters of the first ANN-model.
\end{itemize}\medskip

@@

\subsection{Test system}
The measurements were done on the supercomputer Mistral (phase\,1 system) of the DKRZ (Deutsches Klimarechenzentrum)\footnote{\url{http://www.vi4io.org/hpsl/2016/de/dkrz/start}}.
The phase\,1 system of Mistral operates with the parallel distributed file system Lustre (Lustre 2.5, Seagate's edition) and consists of over 1500 computing nodes and 30 Petabyte storage.
Each computing node consists of two E5-2680v3 with a clock rate of 2.5\,GHz and 30\,MiB L3 Cache.
Measurements were done during daily operation, typical fluctuations in workload may have influenced the execution of benchmarks.

@@

\subsection{Analysis of measurements}
\label{sec:measurements}
The measured access times for the four different cases are summarized in Table\,\ref{overview}.
\tab{overview}{Overview of file access times for the different cases.}{
	\centering
	\scriptsize
	\begin{tabular}{|r|r|r|r|r|r|r|}\hline%
		Case & Min. value  & 1. Quartile & Median & Arith. mean & 3. Quartile & Max. value \\\hline\hline
		\csvreader[late after line=\\\hline]%
		{src/data_summary.csv}{Attribut=\Attribut,Min=\Min,Quartil1=\L, Median = \Median, Mittel = \Mittel,Quartil3 = \Q, Max = \Max, Korrelation = \Korrelation}%
		{\Attribut & \Min & \L & \Median & \Mittel & \Q & \Max}%
	\end{tabular}
}\\
As expected, sequential file accesses are on average much faster than random accesses especially for the reading case, also random read are slower than writes.\\
In Figure \ref{correlations}, we consider the correlations between file access parameters and the resulting access time, a strong correlation can be exploited for access time prediction.
\tab{correlations}{Correlations between file access parameters and access time.}{
	\centering
	\scriptsize
	\begin{tabular}{|r|r|r|}\hline%
		Attribute & Sequential access pattern & Random access pattern\\\hline\hline
		\csvreader[late after line=\\\hline]%
		{src/data_summary_korrelationen.csv}{Attribut=\Attribut,seq = \seq, rnd = \rnd}%
		{\Attribut & \seq &\rnd}%
	\end{tabular}
}\\
Note that sequential access leads to a constant delta-offset of $0$.
Sequential file accesses have with 97.3\% a clear correlation to access time, this is with only 32.91\% to a lesser degree still true for the random accesses.
This is why linear regression can be expected to have acceptable results for access time prediction, especially for the sequential access case.

@@

To study the distribution of measured file access times we display them as follows:
All measurements with equal access size have the same color, access sizes of file accesses are increasing in the graphs from left to right.
The resulting graphs can be seen in the Figures \ref{read_seq} to \ref{write_rnd}.

@@

%\fig{width=.47\textwidth}{src/plot_SizeSorted_log_read_seq2.png}{Measurements of sequential reading file accesses.}
%\fig{width=.47\textwidth}{src/plot_SizeSorted_log_write_seq.png}{Measurements of sequential writing file accesses.}
%\fig{width=.47\textwidth}{src/plot_SizeSorted_log_read_rnd.png}{Measurements of random reading file accesses.}
%\fig{width=.47\textwidth}{src/plot_SizeSorted_log_write_rnd.png}{Measurements of random writing file accesses.}
For readability, the slowest and fastest 1\% of measurements are purged and only every 25th measurement is plot in the figures.\\
The correlation between access size and access time can be easily identified.
As described in chapter \ref{sec:path_for_pred} the split of measurements with identical access parameter values, that occurs in particular in the cases of sequential file accesses, can be explained with a processing in the storage system along different I/O-paths.
It is worth mentioning that the groups of access time for different access sizes are partially at the same range of magnitude.
This indicates as well that the splits are caused by varying I/O-paths as one I/O-path has a typical access time, which only changes slightly for different access sizes.\\
The phenomenon of these splits clarifies why despite of the strong correlation of access size to access time a linear model might produce sub-optimal predictions. It's difficult to predict which I/O-path will be used for a measurement as they differentiate for identical access parameter values.
In Figure~\ref{read_rnd}, the importance of knowledge about I/O-paths for the  access time prediction can be seen, for example file accesses of only 1\,B can be processed as slow as accesses with 16\,MiB if the longest I/O-path is used.\\
For writing file accesses the splits only occur occasionally and are less drastic; these measurements might be well represented with a linear model.

@@

\begin{figure}[!h]
	\centering
	% TODO gibt es das Bild noch so, dass man die Error classes hier mit der selben farbe wie in Fig 10 sehen kann?
	% Muss aber nicht zwingend sein.
	\begin{minipage}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{src/plot_rnd_linreg_Size.png}
		\caption{Predicted access times of the linear regression model as blue points.}
		\label{linreg}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{src/linreg_error_clustering_rnd_all.png}
		\caption{Residues of the linear regression model, colors are corresponding to error classes.}
		\label{clustering}
	\end{minipage}
\end{figure}

@@

The computed error classes can be examined in further detail with the data from Table\,\ref{tab:error_classes}.
Ideally for a good correlation of I/O-paths to error classes, each error class should represent one particular value of data throughput that would be characteristic for the slowest storage medium occurring in the I/O-path. 
This is mostly true for the error classes. The classes 4, 5 and 6, however, have with $5.5\cdot10^7$, $6.1\cdot10^7$ and $5.8\cdot10^7$ very similar average throughput values, which might be a hint for too many error classes.\\
\tab{tab:error_classes}{Characteristics of computed error classes for random file accesses in detail.}{
	\centering
	\scriptsize
	\begin{tabular}{|r|r|r|r|r|r|r|r|}\hline%
		& \multicolumn{3}{|c|}{averaged values} & \multicolumn{3}{|c|}{prediction error} & \multicolumn{1}{r|}{} \\ \hline
		Error class & Throughput (B/s) & Access size (B) & Access time (s) & Min (s) & Avg (s) & Max (s) & \# of members\\\hline\hline
		\csvreader[late after line=\\\hline]%
		{src/rnd_linreg_classes_on_seq.csv}{tp=\tp,idx = \idx,rndcount=\rndcount,minerror=\minerror, meanerror = \meanerror, maxerror = \maxerror, seqcount = \seqcount, size = \size, duration = \duration}%
		{\idx & \tp & \size & \duration & \minerror & \meanerror & \maxerror & \rndcount}%
	\end{tabular}
}\\
The vast majority of measurements were assigned to error class 3, which is the error class that covers the range around a prediction error of 0.
Therefore most access times can be predicted sufficiently by the linear regression model.\\
One sign for correct approximation of I/O-paths is it when error classes exist which have equal access sizes but different access times.
In such cases a differentiation of measurements with equal access parameters takes place.
This occurs for example with the error classes 4 and 8. Both classes have an average access size of $1.2\cdot10^6$ Bytes, but with 0.0143 and 0.2741 seconds very different average access times.
The error classes of sequential file accesses are display similar characteristics as the analyzed random file accesses.
However, since sequential accesses are not able to stress all different I/O path, this result is more interesting.

\subsection{Prediction of file accesses}
After we analyzed the benchmarks and error classes, we study the results of our models.
To investigate the accuracy of the models and investigate overfitting, we split the available data into a training set and a test set. 
The test set consist of all measurements that weren't used for training.
The parameters for the ANN-models are varied to explore appropriate configurations; they are trained with a wide span of different parameter values for the number of hidden layers and number of neurons per layer.
They had 1000 randomly chosen measurements for each pair of access size and access type as training set. % TODO stimmt das 1000 pro Konfiguration? Ja, das stimmt :)
%
% TODO: Wie war das für die großen Zugriffe e.g. 2 MiB, wo keine 10000 Operationen vorlagen?
In total that are 34\,000 measurements which should be a significant amount of data to avoid a situation in which we aren't able to train a model due to a lack of information about the system.

>>>>

%\classify{MSC?}
\author{Jan Fabian Schmid\footnote{\label{susu}Universität Hamburg, E-Mail:\url{2schmid@informatik.uni-hamburg.de}} \and Julian M. Kunkel\footnote{German Climate Computing Center (DKRZ)}}

@@

\keywords{file system, performance, modeling I/O, artificial neural networks}
\end{abstract}

@@

Because of the strong linear correlation between access time and access size, the problem seems to fit linear models.
However, our results show that the relation of file access parameters to access time is not sufficiently represented by linear models.
ANNs achieve significantly better results than linear models.
Our analysis suggests that the I/O-path used by the storage system significantly influences the file access time.
Therefore, it becomes key for a good model of access times to derive knowledge about I/O-paths.
Unfortunately, I/O-paths are difficult to deal with, as it is unknown which path was used for a file access.

This paper is organized as follows:
Related work is provided in Section\,\ref{sec:related}.
In Section\,\ref{modeling_access_times}, we explain our analysis of file access times. 
Firstly, we develop a simplified model of the I/O-path. 
Next, a method for approximating I/O-paths with derived classes from the error of a prediction model is proposed. 
At last, we introduce a set of models for access time prediction.
Afterwards, in Section\,\ref{evaluation}, we evaluate our analysis of the storage system by measuring access times in various test series, studying the obtained access times and the predictions of our models to them.
Finally, Section\,\ref{conclusion} summarizes the paper and suggests future work.

@@

Generally, storage systems are modeled in two different ways for access time prediction; either using white-box modeling or black-box modeling~\cite{Crume:2013:FML:2538542.2538561}.
\begin{itemize}
	\item \textbf{White-box modeling}: The storage system itself is simulated. 
	Details of hardware components like rotation speed of the magnetic disk in a hard drive are considered. 
	The processing of a file access can then be simulated in the model and the resulting access time is then used as prediction for the actual system.
	Processing and resulting performance can be analyzed in detail on the model.
	\item \textbf{Black-box modeling}: The model abstracts from the real storage system. 
	System performance is approximated without considering the causes. %Zustandekommen
	This procedure corresponds to an emulation of a storage system.
	In contrast to the white-box model, processing of file accesses can't be analyzed on the model itself.
\end{itemize}
These two ways of modeling are fundamentally different and have to be differentiated.

\subsection{White-box modeling versus black-box modeling}
For the in-depth analysis of reasoning for behavior of a storage system, a white-box-model is desirable.
On the one hand, the modeled system is represented in the model and, thus, can be examined. 
On the other hand, these models can be very precise if modeled correctly \cite{Ruemmler94anintroduction}.
There are important limitations of white-box modeling, however. 
For every system an individual model has to be created and a model becomes quite intricate for single hard drives already \cite{Crume:2013:FML:2538542.2538561}.
To approach the complexity of white-box modeling, Ruemmler and Wilkes analyzed the relevance of different hard drive components for the model deviation, to save effort for insignificant parts \cite{Ruemmler94anintroduction}.
However, white-box modeling is usually used for simple systems like a single hard drive. For these hard drives white-box-modeling is already very demanding, hence for the complex parallel storage system of a supercomputer it's not a feasible approach \cite{DBLP:conf/npc/ZhangLZJC10}.\\

Application of black-box-modeling is easier and more flexible as it's independent from the individual system.
Stochastic approaches coupled with data mining methods are mostly used for black-box modeling; for example, a combination of regression trees can be used \cite{Dai:2012:SDP:2477169.2477214}, or selective bagging classification and regression trees \cite{DBLP:conf/npc/ZhangLZJC10}.

\subsection{Prediction of I/O-performance with ANNs}
Computability of ANNs was researched by Rojas \cite{Rojas:1996:NNS:235222} and Cybenko \cite{cybenko:mcss}; they demonstrated the  possibility of modeling non-linear systems. 
Cybenko also proved the \textit{universal approximation theorem}, which states that feed-forward networks with sufficient complexity can approximate any continuous functions on compact subsets of $\mathbb{R}^n$.
Therefore, ANNs should be sufficient for predicting I/O-performance as long as there are no contradictions in behavior. 
However, no statement can be made about the necessary structure of the network for this task.\\
Crume et al. developed a method using ANNs which exploits periodic patterns in sequences of file access times \cite{Crume:2013:FML:2538542.2538561}.
A Fourier analysis is used to determine the most important frequencies, which are then used as input information for the ANNs.
In a following publication \cite{crumelatent} they move away from Fourier analysis, but then use additional sine waves as input for the ANNs.
This seems to be a promising approach for predicting access times of single hard drives, where the rotational characteristics of the magnetic disk are an essential consideration. 
It is, however, difficult to extrapolate behavior from a single disk to a distributed storage system consisting of thousands of disks that utilizes several optimization mechanisms.

\subsection{I/O-performance prediction in HPC}
Performance analysis in HPC is an important task to examine and improve system efficiency.
For instance Liu et al. simulated scheduling algorithms for research \cite{liu2011towards}. 
The simulation used the white-box modeling tool DiskSim for prediction of occurring file accesses \cite{Bucy08thedisksim}.\\
There are a few simulators for parallel storage systems, for example, CODES\,\cite{cope2011codes} utilizes a scalable infrastructure to investigate relevant research issues such as the importance of burst-buffers.
PIOSimHD is a simulator able to replay (MPI) application traces on a generic I/O system \cite{kunkel2013simulating}.

@@

\section{Modeling I/O Access Times}
\label{modeling_access_times}
\subsection{Characteristics of the Data}
We used a synthetic benchmark in which identical measurements for random or sequential I/O are performed and the access time is measured for each operation individually.
The resulting timings are stored together with file access parameters in a file that is then used to build and validate the models.
This approach allows for an in-depth analysis of system behavior for different use cases from the perspective of a single client.
The stored parameters are:
\begin{itemize}
%	\item \textbf{File-ID}: Identifies the invoked file. Brauchen wir ja nicht für die Analyse
	\item \textbf{Access size}: Number of bytes to read or write.
	\item \textbf{Access type}: Differentiates reading and writing.
\end{itemize}
Directly measurable or derivable attributes are:
\begin{itemize}
	\item \textbf{Offset}: Distance of file beginning to the starting point of access.
	\item \textbf{Delta-Offset}: Can be calculated for file access $i$ as Offset[$i$] - (Offset[i-1] + access size[i-1]).
	\item \textbf{Access time}: Time for performing the I/O.
\end{itemize}
Knowledge of internal aspects of the system about the current system utilization or about the storage media that have to be addressed are not used by the model.

\subsection{Model of the I/O-path}
The internal processing of a file access in the storage system can be viewed using the I/O-path, which is the path from the invoking processor to the storage medium that contains/will store the data (larger accesses may be distributed across multiple servers). 
The data read/written is passed through the levels of the memory hierarchy.
The resulting access time depends on the depth of this I/O-path, because with increasing distance the storage media along the path is slower.
While the first levels in memory hierarchy (CPU caches) are the fastest volatile storage, the main memory is already an order of magnitude slower; the same applies for the step into the parallel storage system, that is deployed as servers connected via network to the computer nodes.
%The trade-off between memory capacity and performance can be visualized with the memory hierarchy as in figure %\ref{src/2000px-ComputerMemoryHierarchy.png}.
%\fig{width=.7\textwidth}{src/2000px-ComputerMemoryHierarchy.png}{Memory hierarchy.}
Reading file accesses are more severely affected by varying depths of I/O-paths than writing file accesses, which can be deferred using write-behind and propagated lazily to the disk drives.

@@

At a first consideration, one could come up with the idea of simply clustering the data seen in Figure~\ref{src/plot_SizeSorted_log_read_seq.png}, because the groups of measurements with varying I/O-paths might be correctly differentiated by it. 
This, however, is not a good approach. One would have to cluster for every set of measurements with identical access parameters individually as access time is strongly dependent on the access size. 
In a real application scenario, a large number of different access parameters with only very few instances occur, which makes this approach unfeasible.\\
Another idea might be to use the difference of an arbitrary value to the measured file access time as parameter for the clustering. 
Such a value would not be appropriate for all magnitudes of access time occurring in the data.
I/O-paths that differ by little might not be found with this approach, because the gap between them doesn't matter at the scale of the chosen value.

Our approach uses the residues of a model like linear regression that can't differentiate between measurements with equal access parameter values.
Basically, the linear regression predicts an average access time for the measurements of any given size.
The deviation of an individual measurement for these access parameters to the predicted average value is then characteristic for the I/O-path.
Clustering the residues with a k-Means algorithm allows us then to assign each measurement to a specific cluster, which is the approximation of its I/O-path.
We call the retrieved clusters \textbf{error classes}.
We use the absolute error of linear regression as clustering parameter so that one cluster can, for example, represent an positive deviation of 1 millisecond.
Ideally, it might be the case that file accesses independent of the access parameters that took 1 millisecond longer than usual are processed on the same I/O-path.
In that case error classes are directly corresponding to I/O-paths. 
However, this has to be investigated.

@@

\subsection{Models}
Access times of file accesses can be predicted using various models.
While we evaluated several models for this paper, only a few relevant models are described; more models are described in \cite{VVEIHUDVVN15}. 
% TODO: Soll ich meine Arbeit noch woanders zitieren?
Each model seeks to correlate its known file access parameters to the corresponding access time. 
Linear regression is used as a baseline model with a simple mapping of access size to access time.
Additionally, three models with different input information utilizing ANNs are used:
\begin{itemize}
	\item The most simple ANN-model only received information about access sizes, delta-offsets and access types as input.
	\item A second ANN-model, called \textbf{ema-model}, receives additionally to the parameters of the previous model,  information about the past data throughputs of the system. This can be used to exploit time dependencies of the I/O-performance.
	\item The third ANN-model, called \textbf{error-class-model}, received error classes in addition to the access parameters of the first ANN-model.
\end{itemize}\medskip

@@

\subsection{Test system}
The measurements were done on the supercomputer Mistral (phase\,1 system) of the DKRZ (Deutsches Klimarechenzentrum)\footnote{\url{http://www.vi4io.org/hpsl/2016/de/dkrz/start}}.
The phase\,1 system of Mistral operates with the parallel distributed file system Lustre (Lustre 2.5, Seagate's edition) and consists of over 1500 compute nodes and 30 Petabyte storage.
Each compute node consists of two E5-2680v3 with a clock rate of 2.5\,GHz and 30\,MiB L3 Cache.
Measurements were done during daily operation, typical fluctuations in workload may have influenced the execution of benchmarks.

@@

\subsection{Analysis of measurements}
\label{sec:measurements}
The measured access times for the four different cases are summarized in Table\,\ref{overview}.
\tab{overview}{Overview of file access times for the different cases.}{
	\centering
	\scriptsize
	\begin{tabular}{|r|r|r|r|r|r|r|}\hline%
		Case & Min. value  & 1. Quartile & Median & Arith. mean & 3. Quartile & Max. value \\\hline\hline
		\csvreader[late after line=\\\hline]%
		{src/data_summary.csv}{Attribut=\Attribut,Min=\Min,Quartil1=\L, Median = \Median, Mittel = \Mittel,Quartil3 = \Q, Max = \Max, Korrelation = \Korrelation}%
		{\Attribut & \Min & \L & \Median & \Mittel & \Q & \Max}%
	\end{tabular}
}\\
As expected, sequential file accesses are on average much faster than random accesses especially for the reading case, also random reads are slower than writes.\\
In Figure \ref{correlations}, we consider the correlations between file access parameters and the resulting access time, a strong correlation can be exploited for access time prediction.
\tab{correlations}{Correlations between file access parameters and access time.}{
	\centering
	\scriptsize
	\begin{tabular}{|r|r|r|}\hline%
		Attribute & Sequential access pattern & Random access pattern\\\hline\hline
		\csvreader[late after line=\\\hline]%
		{src/data_summary_korrelationen.csv}{Attribut=\Attribut,seq = \seq, rnd = \rnd}%
		{\Attribut & \seq &\rnd}%
	\end{tabular}
}\\
Note that sequential access leads to a constant delta-offset of $0$.
Sequential file accesses have with 97.3\% a clear correlation to access time, this is with only 32.91\% to a lesser degree still true for the random accesses.
This is why linear regression can be expected to have acceptable results for access time prediction, especially for the sequential access case.

@@

To study the distribution of measured file access times we display them as follows:
All measurements with equal access size have the same color, access sizes of file accesses are increasing in the graphs from left to right.
The resulting graphs can be seen in the Figures \ref{read_seq} to \ref{write_rnd}. Note that they use logarithmic scale.

@@

%\fig{width=.47\textwidth}{src/plot_SizeSorted_log_read_seq2.png}{Measurements of sequential reading file accesses.}
%\fig{width=.47\textwidth}{src/plot_SizeSorted_log_write_seq.png}{Measurements of sequential writing file accesses.}
%\fig{width=.47\textwidth}{src/plot_SizeSorted_log_read_rnd.png}{Measurements of random reading file accesses.}
%\fig{width=.47\textwidth}{src/plot_SizeSorted_log_write_rnd.png}{Measurements of random writing file accesses.}
For readability, the slowest and fastest 1\% of measurements are purged and only every 25th measurement is plot in the figures.\\
The correlation between access size and access time can be easily identified.
As described in chapter \ref{sec:path_for_pred} the split of measurements with identical access parameter values, that occurs in particular in the cases of sequential file accesses, can be explained with a processing in the storage system along different I/O-paths.
It is worth mentioning that the groups of access time for different access sizes are partially at the same range of magnitude.
This indicates as well that the splits are caused by varying I/O-paths as one I/O-path has a typical access time, which only changes slightly for different access sizes.\\
The phenomenon of these splits clarifies why despite of the strong correlation of access size to access time a linear model might produce sub-optimal predictions. 
Even for identical access parameter, it is difficult to predict which I/O-path will be used as the storage system decides the I/O path based on the system state (e.g., is the data available in the client's cache).
In Figure~\ref{read_rnd}, the importance of knowledge about I/O-paths for the  access time prediction can be seen, for example file accesses of only 1\,B can be processed as slow as accesses with 16\,MiB if the longest I/O-path is used.\\
For writing file accesses the splits only occur occasionally and are less drastic; these measurements might be well represented with a linear model.

@@

\begin{figure}[!h]
	\centering
	% TODO gibt es das Bild noch so, dass man die Error classes hier mit der selben farbe wie in Fig 10 sehen kann?
	% Muss aber nicht zwingend sein.
	%TODO nee, das gibt es leider nicht.	
	\begin{minipage}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{src/plot_rnd_linreg_Size.png}
		\caption{Predicted access times of the linear regression model as blue points.}
		\label{linreg}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{src/linreg_error_clustering_rnd_all.png}
		\caption{Residues of the linear regression model, colors are corresponding to error classes.}
		\label{clustering}
	\end{minipage}
\end{figure}

@@

The computed error classes can be examined in further detail with the data from Table\,\ref{tab:error_classes}.
Ideally for a good correlation of I/O-paths to error classes, each error class should represent one particular value of data throughput that would be characteristic for the slowest storage medium occurring in the I/O-path. 
This is mostly true for the error classes. The classes 4, 5 and 6, however, have with $5.5\cdot10^7$, $6.1\cdot10^7$ and $5.8\cdot10^7$ very similar average throughput values, which might be a hint for too many error classes.\\
\tab{tab:error_classes}{Characteristics of computed error classes for random file accesses in detail.}{
	\centering
	\scriptsize
	\begin{tabular}{|r|r|r|r|r|r|r|r|}\hline%
		& \multicolumn{3}{|c|}{averaged values} & \multicolumn{3}{|c|}{prediction error} & \multicolumn{1}{r|}{} \\ \hline
		Error class & Throughput (B/s) & Access size (B) & Access time (s) & Min (s) & Avg (s) & Max (s) & \# of members\\\hline\hline
		\csvreader[late after line=\\\hline]%
		{src/rnd_linreg_classes_on_seq.csv}{tp=\tp,idx = \idx,rndcount=\rndcount,minerror=\minerror, meanerror = \meanerror, maxerror = \maxerror, seqcount = \seqcount, size = \size, duration = \duration}%
		{\idx & \tp & \size & \duration & \minerror & \meanerror & \maxerror & \rndcount}%
	\end{tabular}
}\\
The vast majority of measurements were assigned to error class 3, which is the error class that covers the range around a prediction error of 0.
Therefore, most access times can be predicted sufficiently by the linear regression model.\\
One sign for correct approximation of I/O-paths is it when error classes exist which have equal access sizes but different access times.
In such cases a differentiation of measurements with equal access parameters takes place.
This occurs for example with the error classes 4 and 8. Both classes have an average access size of $1.2\cdot10^6$ Bytes, but with 0.0143 and 0.2741 seconds very different average access times.
The error classes of sequential file accesses are display similar characteristics as the analyzed random file accesses.
However, since sequential accesses are not able to stress all different I/O path, this result is more interesting.

\subsection{Prediction of file accesses}
Next, we study the results of our models.
To investigate the accuracy of the models and investigate overfitting, we split the available data into a training set and a test set. 
The test set consist of all measurements that weren't used for training.
The parameters for the ANN-models are varied to explore appropriate configurations; they are trained with a wide span of different parameter values for the number of hidden layers and number of neurons per layer.
They had 1000 randomly chosen measurements for each pair of access size and access type as training set. % TODO stimmt das 1000 pro Konfiguration? Ja, das stimmt :)
%
% TODO: Wie war das für die großen Zugriffe e.g. 2 MiB, wo keine 1000 Operationen vorlagen?

%TODO: Du hast Recht, das ist ein Problem für den Fall mit 16MiB, da gibt es nur 640 Messungen, die sind dann alle zum Trainieren genutzt wurden. Sind dann aber auch nicht für die Berechnung der Residuen genutzt wurden. Für alle anderen Zugriffsgrößen gibt es mehr als 1000 Messungen.
In total that are 34\,000 measurements which should be a significant amount of data to avoid the situation in which too few data points are available to train the system.
